---
title: "Buoyant Weight Analysis"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/fragGrowthExperiment.git){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from 3D models of micro-fragmented coral colonies and larger frags to assess differences in growth rates between microfragments and natural growth. 

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "dplyr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "vegan", "googledrive", "gdata", "readxl", "DescTools","patchwork","reshape2", "lme4", "emmeans", "MuMIn", "sjstats", "lmerTest")




```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 



```{r, loadingData, include = TRUE}

# drive_download(
# as_id("1IK7_sd_BBPcuxHgUC8sfFOp1oqGOb0fE"),
# path = "../data/bwDataset.xlsx",
# overwrite = TRUE)

bw <- read_excel("../data/bwDataset.xlsx")
bw

```
# Filtering Data
***
## Now we will perform a number of filters and data manipulations using multiple functions from the package `dplyr` to distill our data into a more digestable format. First, we want to combine two columns **Species** and **Genotype** into one column called **Geno**. We are doing this using hte `paste0` function. Then we are filtering our dataset to only include the columns we want using the `select()` function from the package `dplyr`. 

```{r, dataManipulation, include = TRUE}

bw$genotype <- as.character(bw$genotype)

bw$geno <- paste0(bw$species, bw$genotype)

bw$species <- as.factor(bw$species)

bwBlanks <- bw %>% 
  select("date", "timePoint", "species","geno", "trial", "ID", "dryWeight", "n") %>% 
  dplyr::filter(ID %in% c("plugs", "tile"))
  
bwClean <- bw %>% 
  select("date", "timePoint", "species","geno", "trial", "ID", "dryWeight", "n") %>% 
  dplyr::filter(species %in% c("pc", "ps", "dl"))

bwClean <- rbind(bwClean, bwBlanks)


bwClean$timePoint <- as.factor(bwClean$timePoint)

bwClean$species <- as.factor(bwClean$species)

bwClean$geno <- as.factor(bwClean$geno)

bwClean$trial <- as.factor(bwClean$trial)

bwClean$ID <- as.factor(bwClean$ID)


head(bwClean)




```

***
## Here I am going to separate our data into only the time points we want. In this case, it is first (T0), middle (T3) and final (T6). 

```{r, sepTimePoints, include = TRUE}


bwTime <- bwClean %>% 
  dplyr::select("date", "timePoint", "species","geno", "trial", "ID", "dryWeight", "n") %>% 
  dplyr::filter(timePoint %in% c("0", "3", "6"))





```





## Here I am going to make the corrections from the blank plugs and tiles and incorporate them into our dry weight measurements. We had a blank tile and 5 blank plugs in each raceway to account for algae growth and fouling oraganisms (mainly tube worms) fouling our corals. 

```{r, blankCorrection, include = TRUE}


bwTiles <- bwTime %>% 
  select("date", "timePoint", "species", "geno", "trial", "ID", "dryWeight", "n") %>% 
  dplyr::filter(ID %in% "tile")




meanTiles <- mean(bwTiles$dryWeight)



#BELOW MIGHT NOT BE NECESSARY, NOT SURE

# bwTileChange <- dcast(bwTiles, trial+ID~timePoint) 
# 
# bwTileChange <- dplyr::rename(bwTileChange, "T0" = "0", "T3" = "3", "T6" = "6")
# bwTileChange
# 
# bwTileChange$percentChangeFirstInterval <- ((bwTileChange$T3-bwTileChange$T0)/bwTileChange$T0)*100
# bwTileChange$percentChangeSecondInterval <- ((bwTileChange$T6-bwTileChange$T3)/bwTileChange$T3)*100
# bwTileChange$percentChangeTotalInterval <- ((bwTileChange$T6-bwTileChange$T0)/bwTileChange$T0)*100
# 



bwPlugs <- bwTime %>% 
  select("date", "timePoint", "species", "geno", "trial", "ID", "dryWeight", "n") %>% 
  dplyr::filter(ID %in% "plugs")






bwPlugs$avgPlug <- bwPlugs$dryWeight/bwPlugs$n
meanPlug <- mean(bwPlugs$avgPlug)





```


### Now we are going to correct our weights by removing the weight of the plugs and the tiles from the frags and whole colonies, respectively. First, we need to multiply the weight of the plug by the n of each rack. E.g. If a rack has 26 plugs on it we need to multiply our `meanPlug` number by 26 to account for that rack. Then we will subtract that from our dryWeigh. Similarly, we are taking the average of our blank tile weight (`meanTiles`) and subtracting that from our whole colonies. 


```{r, dryWeightCorrection, include = TRUE}


bwCleanFrags <- bwTime %>% 
  dplyr::filter(ID %in% "frag")

meanPlug <- rep(meanPlug, length(bwCleanFrags$n))
bwCleanFrags <- cbind(bwCleanFrags, as_tibble(meanPlug))
bwCleanFrags$correctionFactor <- bwCleanFrags$n * bwCleanFrags$value

bwCleanFrags$correctedDryWeight <- bwCleanFrags$dryWeight - bwCleanFrags$correctionFactor


bwCleanWild <- bwTime %>% 
  dplyr::filter(ID %in% "wild")

bwCleanWild$correctedDryWeight <- bwCleanWild$dryWeight - meanTiles



```




## Now I want to combine our data back to one data frame with the corrected dry weights of the frags vs wilds.

```{r, splittingData, include = TRUE}

bwCleanFrags <- bwCleanFrags %>% 
  select(date, timePoint, species, geno, trial, ID, dryWeight, n, correctedDryWeight)
bwClean1 <- rbind(bwCleanFrags, bwCleanWild)

totalFrag <- bwClean1 %>%
  select(timePoint, species, ID, geno,trial, correctedDryWeight) %>%
  dplyr::filter(ID %in% 'frag') %>%
  dplyr::group_by(timePoint, geno, species, ID) %>%
  dplyr::summarise(sum(correctedDryWeight)) %>% 
  dplyr::rename("correctedDryWeight" = "sum(correctedDryWeight)") %>% 
  ungroup()
totalWhole <- bwClean1 %>% 
  select(timePoint, species, ID, geno,correctedDryWeight) %>%
  dplyr::filter(ID %in% 'wild')


bwClean <- rbind(totalFrag, totalWhole)


```

# Calculating Percent Change
***
Because the wilds and frags started with wildly different amounts of skeleton we want to calculate rate of change to try not only to normalize the data, but also to investigate which treatment is producing more skeleton. 

```{r, percentChange, include = TRUE}

bwChange <- dcast(bwClean, geno+species+ID~timePoint) 


bwChange$`T3-T0` <- ((bwChange$`3`-bwChange$`0`)/bwChange$`0`)*100
bwChange$`T6-T3` <- ((bwChange$`6`-bwChange$`3`)/bwChange$`3`)*100
bwChange$`T6-T0` <- ((bwChange$`6`-bwChange$`0`)/bwChange$`0`)*100

# bwChange1 <- bwChange %>% 
#             select(geno, species, ID, `0`, `3`, `6`)
bwChange2 <- bwChange %>% 
            select(geno, species, ID, `T3-T0`, `T6-T3`, `T6-T0`)

# bwChangeMelt1 <- melt(bwChange1, id = c(1,2,3))
# bwChangeMelt1 <- bwChangeMelt1 %>% 
#                 dplyr::rename("timePoint" = "variable") %>% 
#                 dplyr::rename("correctedDryWeight" = "value")    

bwChangeMelt2 <- melt(bwChange2, id = c(1,2,3))
bwChangeMelt2 <- bwChangeMelt2 %>% 
                dplyr::rename("timeSeries" = "variable") %>% 
                dplyr::rename("percentChange" = "value")    

bwChange <- bwChangeMelt2
# bwChange <- dplyr::left_join(bwChangeMelt1, bwChangeMelt2, by = c("geno", "species", "ID","timePoint")) %>% 
#   select(geno, species, ID, timeSeries, percentChange)
bwChange$timeSeries <- as.factor(bwChange$timeSeries)


bwChange <- bwChange %>% 
        filter(timeSeries == "T6-T0")


```




# Checking for Normality
***
## I am checking to see if the data fits a normal distribution

```{r, normalityCheck, include = TRUE}

shapiro.test(bwClean$correctedDryWeight)
hist(bwClean$correctedDryWeight)


shapiro.test(bwChange$percentChange)
hist(bwChange$percentChange)

```
# Statistical Tests
***
## The data is not normally distributed and transformations failed to achieve normality. 
```{r, stats, include = TRUE}


totalLM <- lmer(correctedDryWeight ~ species*ID*timePoint + (1 | geno), bwClean)
totalLManova <- anova(totalLM)
totalLManova


eta_sq(totalLM, partial = TRUE)
r.squaredGLMM(totalLM)

#timePoint, and the interaction between species:ID are significant 
emmeans(totalLM, list(pairwise ~ timePoint*ID), adjust = 'bonferroni')
emmeans(totalLM, list(pairwise ~ timePoint*ID), adjust = 'bonferroni')

changeLM <- lmer(percentChange ~ species*ID*timeSeries + (1 | geno), bwChange)
changeLManova <- anova(changeLM)
changeLManova


eta_sq(changeLM, partial = TRUE)
r.squaredGLMM(changeLM)
emmeans(changeLM, list(pairwise ~ ID*timeSeries), adjust = 'bonferroni')

cleanFrag <- bwClean %>% 
  filter(ID %in% "frag")

cleanWhole <- bwClean %>% 
  filter(ID %in% "wild")
mean(totalFrag$correctedDryWeight)

mean(cleanWhole$correctedDryWeight)


cleanChangeFrag <- bwChange %>% 
  filter(ID %in% "frag")
cleanChangeWhole <- bwChange %>% 
  filter(ID %in% "wild")
mean(cleanChangeFrag$percentChange)
mean(cleanChangeWhole$percentChange)


```



# Visualizing Data
***
## I now want to visualize our data using the package `ggplot2`. But first I am going to subset the larger dataset to remove the blanks and plugs from the dataset. 


```{r, meanPercentChangeVisualization, include = TRUE}
# could do confidence interval 

se <- function(x) sqrt(var(x) / length(x))

totalBWse <- bwClean %>%
  dplyr::select(timePoint, geno, species, ID, correctedDryWeight) %>% 
  dplyr::group_by(timePoint, species, ID) %>% 
  dplyr::summarise(
    meanBW = mean(correctedDryWeight),
    sd = sd(correctedDryWeight),
    n = n(),
    se = sd/sqrt(n))

totalChangeSe <- bwChange %>% 
  dplyr::select(timeSeries, geno, species, ID, percentChange) %>% 
  dplyr::group_by(timeSeries, species, ID) %>% 
  dplyr::summarise(
    meanPC = mean(percentChange),
    sd = sd(percentChange),
    n = n(),
    se = sd/sqrt(n))


bwPercentChange.1 <- ggplot(totalChangeSe, aes(x = ID, y = meanPC, fill = timeSeries)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~species)+
    ggtitle("Buoyant Weight Percent Growth of D. labyrinthiformis, P. strigosa, and P. clivosa") + scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Whole"))+scale_fill_manual(name="Time Point",values = c( "#EAA65D","#0D3B66"))+    
    xlab("Treatment") + ylab("Percent Change (%)")+
  geom_errorbar( aes(x=ID, ymin=meanPC-se, ymax=meanPC+se), position = position_dodge()) +
   # ylim(0,100)+    
    xlab("Treatment") + ylab("Percent Change (%)")
bwPercentChange.1

bwPercentChange<- bwPercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_blank(), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 35, color="black"), 
  axis.text.x=element_text(size = 31, color="black"), 
  axis.text.y=element_text(size = 31, color="black"),
  legend.position = "right")

bwPercentChange

ggsave("../figures/bwPercentChange.png", plot = bwPercentChange, height = 15, width = 20, units = 'in',dpi = 600)

```

#Here we are creating a figure for the mean T0, T3, and T6 surface area of each species. Included in this figure are error bars derived from the standard deviation divided by the square root of the sample size (n)

```{r, meanSAgrowthVisualization, include = TRUE}



bwGrowth.1 <- ggplot(data=totalBWse, aes(x = ID, y = meanBW, fill = timePoint)) +
  geom_bar(stat = "identity", position = position_dodge(), alpha = 0.75)+
  geom_errorbar( aes(x=ID, ymin=meanBW-se, ymax=meanBW+se), position = position_dodge())+ 
  facet_wrap(~species)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("BuoyantWeight (g)"),fill='Time Point') + 
  ggtitle("Mean Skeletal Growth of D. labyrinthiformis, P. strigosa, and P. clivosa") +
  scale_fill_discrete(labels=c('T0','T3','T6'))+
  scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Whole"))+
  scale_fill_manual(values = c("#0053A6", "#D3F292", "#5CADFF"))

bwGrowth.1

bwGrowth <- bwGrowth.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_blank(), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 35, color="black"), 
  axis.text.x=element_text(size = 31, color="black"), 
  axis.text.y=element_text(size = 31, color="black"),
  legend.position = "right")


bwGrowth

ggsave("../figures/bwGrowth.png", plot = bwGrowth, height = 15, width = 20, units = 'in',dpi = 600)

```




```{r, PERMANOVA, include = TRUE}


plotFrags <- bwChangeAov %>% 
  filter(ID %in% 'frag')
plotWilds <-bwChangeAov %>% 
  filter(ID %in% 'wild')


percentChangeFragPlotAovBW <- ggplot(plotFrags, aes(x = geno, y = percentChange))+
  geom_bar(stat = 'identity', position = 'dodge', fill = 'darkblue')+
  ylim(-10, 50)+
  facet_wrap(trial~.)
percentChangeWildPlotAovBW <- ggplot(plotWilds, aes(x = geno, y = percentChange))+
 geom_bar(stat = 'identity', position = 'dodge', fill = 'navy')+
  ylim(-10, 50)


comboPlotAovBW <- percentChangeFragPlotAovBW + percentChangeWildPlotAovBW


ggsave("../figures/Fig2.png", plot = percentChangeFragPlotAov, height = 5, width = 9, units = 'in',dpi = 600)

```

```{r, plotChange, include=TRUE}

percentChangeFragPlotAovBW <- percentChangeFragPlotAovBW + 
  labs(title = "Skeletal Deposition of Microfragments", 
       subtitle = "Graph of Percent Change in Buoyant Weight by Species and Genotype", 
       x= "Coral Species and Genotype", 
       y= "Buoyant Weight Percent Change")

percentChangeFragPlotAovBW <- percentChangeFragPlotAovBW + theme(
  plot.title = element_text(color = "black", size = 20, 
                            face = "bold", hjust = 0.5),
  plot.subtitle = element_text(color = "black", size = 12,
                               hjust = 0.5),
  axis.title.x = element_text(face = "bold"),
  axis.title.y = element_text(face = "bold"),
  axis.text.x = element_text(angle = 45),
)
 
percentChangeWildPlotAovBW <- percentChangeWildPlotAovBW + 
  labs(title = "Skeletal Deposition of Wild Colonies", 
       subtitle = "Graph of Percent Change in Buoyant Weight by Species and Genotype", 
       x= "Coral Species and Genotype", 
       y= "Buoyant Weight Percent Change")

percentChangeWildPlotAovBW <- percentChangeWildPlotAovBW + theme(
  plot.title = element_text(color = "black", size = 20, 
                            face = "bold", hjust = 0.5),
  plot.subtitle = element_text(color = "black", size = 12,
                               hjust = 0.5),
  axis.title.x = element_text(face = "bold"),
  axis.title.y = element_text(face = "bold"),
  axis.text.x = element_text(angle = 45),
)




```




