---
title: "Surface Area Analysis"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/fragGrowthExperiment.git){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from 3D models of micro-fragmented coral colonies and larger frags to assess differences in growth rates between microfragments and natural growth. 

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "vegan", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse", "lme4", 'PMCRMplus', "EnvStats", "emmeans", "MuMIn", "sjstats", "lmerTest")

#IF YOU WANT TO RUN PAIRWISE PERMANOVA REMEMBER TO INSTALL PAIRWISE ADONIS

```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 
```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

dlsa <- read_excel("../data/saDataset.xlsx", sheet = 3)

dlsa


```
## Now I am just tidying up my dataset, I am combining two columns, **species** and **genotype** into a column **geno** using the `paste0` function, cleaning up the raw dataset to only include my species of concern using multiple functions from the package `dplyr`, and changing the class of some variables to factors using the function `as.factor()` for easier manipulation down the road.
```{r, dataManipulation, include = TRUE}


dlsaClean <- dlsa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




dlsaClean$timepoint <- as.factor(dlsaClean$timepoint)

dlsaClean$species <- as.factor(dlsaClean$species)

dlsaClean$id <- as.factor(dlsaClean$id)

dlsaClean$trial <- as.factor(dlsaClean$trial)

dlsaClean$treatment <- as.factor(dlsaClean$treatment)



head(dlsaClean)



```
### Here we are breaking the data apart into frags and wilds, combinging the frags from each trial, putting the data back together, and then calculating percent change of surface area between the first and last timepoint. We are using a number of functions from the package `dplyr` as well as the function `dcast()` to recombine data from the package `reshape2` 
```{r, calculatingPercentChange, include = TRUE }


dlsaFrag <- dlsaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


dlsaWild <- dlsaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


dlsaCombo <- bind_rows(dlsaFrag,dlsaWild)

dlsaFinal <- dlsaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



dlsaChange <- dcast(dlsaFinal, id+treatment~timepoint) 


dlsaChange <- dplyr::rename(dlsaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
dlsaChange
dlsaChange$difference <- (dlsaChange$finalSurfaceArea-dlsaChange$initialSurfaceArea)
dlsaChange$percentChange <- ((dlsaChange$finalSurfaceArea-dlsaChange$initialSurfaceArea)/dlsaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



dlSApercentChange.1 <- ggplot(data=dlsaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) +  scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E")) +     
    xlab("ID") + ylab("Percent Change (%)") + # Set axis labels
    ggtitle("D. lab") +     # Set title
    theme_bw()


dlSApercentChange <- dlSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

dlSApercentChange.1

ggsave("../figures/dlPercentSAgrowth.png", plot = dlSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



dlSAgrowthGenotype.1 <- ggplot(data=dlsaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of DL01, DL02, and DL03") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#3BB0E3", "#eba862", "#12A528"))

dlSAgrowthGenotype.1

dlSAgrowthGenotype <- dlSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


dlSAgrowthGenotype

ggsave("../figures/dlSAgrowth.png", plot = dlSAgrowthGenotype, height = 5, width = 9, units = 'in',dpi = 600)



```

```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

pcsa <- read_excel("../data/saDataset.xlsx", sheet = 4)

pcsa


```

```{r, dataManipulation, include = TRUE}


pcsaClean <- pcsa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




pcsaClean$timepoint <- as.factor(pcsaClean$timepoint)

pcsaClean$species <- as.factor(pcsaClean$species)

pcsaClean$id <- as.factor(pcsaClean$id)

pcsaClean$trial <- as.factor(pcsaClean$trial)

pcsaClean$treatment <- as.factor(pcsaClean$treatment)



head(pcsaClean)



```

```{r, calculatingPercentChange, include = TRUE }


pcsaFrag <- pcsaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


pcsaWild <- pcsaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


pcsaCombo <- bind_rows(pcsaFrag,pcsaWild)

pcsaFinal <- pcsaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



pcsaChange <- dcast(pcsaFinal, id+treatment~timepoint) 


pcsaChange <- dplyr::rename(pcsaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
pcsaChange
pcsaChange$difference <- (pcsaChange$finalSurfaceArea-pcsaChange$initialSurfaceArea)
pcsaChange$percentChange <- ((pcsaChange$finalSurfaceArea-pcsaChange$initialSurfaceArea)/pcsaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



pcSApercentChange.1 <- ggplot(data=pcsaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) + scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E")) +   
    xlab("ID") + ylab("Percent Change (%)") + # Set axis labels
    ggtitle("P. cli") +     # Set title
    theme_bw()

pcSApercentChange <- pcSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

pcSApercentChange

ggsave("../figures/pcPercentSAgrowth.png", plot = pcSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



pcSAgrowthGenotype.1 <- ggplot(data=pcsaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of PC13, PC15, and PC16") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#3BB0E3", "#eba862", "#80f191"))

pcSAgrowthGenotype.1

pcSAgrowthGenotype <- pcSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


pcSAgrowthGenotype

ggsave("../figures/pcSAgrowth.png", plot = pcSAgrowthGenotype, height = 5, width = 9, units = 'in',dpi = 600)



```

```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

pssa <- read_excel("../data/saDataset.xlsx", sheet = 5)

pssa


```

```{r, dataManipulation, include = TRUE}


pssaClean <- pssa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




pssaClean$timepoint <- as.factor(pssaClean$timepoint)

pssaClean$species <- as.factor(pssaClean$species)

pssaClean$id <- as.factor(pssaClean$id)

pssaClean$trial <- as.factor(pssaClean$trial)

pssaClean$treatment <- as.factor(pssaClean$treatment)



head(pssaClean)



```

```{r, calculatingPercentChange, include = TRUE }


pssaFrag <- pssaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


pssaWild <- pssaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


pssaCombo <- bind_rows(pssaFrag,pssaWild)

pssaFinal <- pssaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



pssaChange <- dcast(pssaFinal, id+treatment~timepoint) 


pssaChange <- dplyr::rename(pssaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
pssaChange
pssaChange$difference <- (pssaChange$finalSurfaceArea-pssaChange$initialSurfaceArea)
pssaChange$percentChange <- ((pssaChange$finalSurfaceArea-pssaChange$initialSurfaceArea)/pssaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



psSApercentChange.1 <- ggplot(data=pssaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) +scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E"))+      
    xlab("ID") + ylab("Percent Change (%)") + 
    ggtitle("P. str") +     
    theme_bw()

psSApercentChange <- psSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

psSApercentChange

ggsave("../figures/psPercentSAgrowth.png", plot = psSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



psSAgrowthGenotype.1 <- ggplot(data=pssaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of PS07, PS11, PS12, PS13, and PS14") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#82CDED", "#2B4F2E", "#EAA65D"))

psSAgrowthGenotype.1

psSAgrowthGenotype <- psSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


psSAgrowthGenotype

ggsave("../figures/psSAgrowth.png", plot = psSAgrowthGenotype, height = 10, width = 15, units = 'in',dpi = 600)



```

```{r}
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

percentChange <- ggarrange(dlSApercentChange, pcSApercentChange, psSApercentChange,
                    labels = c("A", "B", "C"),
                    ncol = 2, nrow = 2, common.legend = TRUE, legend = "right")
percentChange <- annotate_figure(
  percentChange,
  top = text_grob("Surface Area Percent Change",
                  color = "black", face = "bold", size = 24))


# ggarrange(
#   dlSApercentChange,                
#   ggarrange(pcSApercentChange, psSApercentChange, ncol = 2, labels = c("B", "C")), 
#   nrow = 2, 
#   labels = "A"      
#   ) 
percentChange

ggsave("../figures/percentChange.png", plot = percentChange, height = 15, width = 20, units = 'in',dpi = 600)

```
```{r}

percentChange.1<-dlSApercentChange+ pcSApercentChange+psSApercentChange+plot_layout(nrow = 2, byrow = FALSE)+plot_annotation(title = 'Surface Area Percent Change',
                  theme = theme(plot.title = element_text(size = 24)))

percentChange.1

ggsave("../figures/percentChange.1.png", plot = percentChange.1, height = 15, width = 20, units = 'in',dpi = 600)
```

Now we will run a paired t-test to determine if the percent change in surface area of the frags varies from the wilds. First, we need to assess normality via the shapiro-wilk test for normal distribution and the Levene's Test to test for homogeneity of variance. If neither is significant, we will conduct a t-test.
```{r}
dlsaFragTest <- dlsaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

dlsaWildTest <- dlsaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

dlsaTest <- left_join(dlsaFragTest, dlsaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(fragpercentChange = percentChange.x) %>%
  dplyr::rename(wildpercentChange = percentChange.y)
```


```{r}
pcsaFragTest <- pcsaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

pcsaWildTest <- pcsaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

pcsaTest <- left_join(pcsaFragTest, pcsaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(pcfragpercentChange = percentChange.x) %>%
  dplyr::rename(pcwildpercentChange = percentChange.y)
```



```{r}
pssaFragTest <- pssaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

pssaWildTest <- pssaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

pssaTest <- left_join(pssaFragTest, pssaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(psfragpercentChange = percentChange.x) %>%
  dplyr::rename(pswildpercentChange = percentChange.y)
```





# Combining Data
***
## We are going to combine all of the data we've parsed out above into two larger data frames. One for the raw surface area numbers `totalSA` and the other for the percent change data `totalChange`. The latter of which requires a bit more intensive data manipulation because we are calculating percent change of two different timepoints, and combining those data. All of this is done using `dplyr` and `reshape2`.

```{r, combiningData, include = TRUE}
# Combining the smaller data sets above into two larger data sets 
totalSA <- bind_rows(dlsaFinal, pcsaFinal, pssaFinal)
totalClean <- bind_rows(dlsaClean, pcsaClean, pssaClean)

# Now we need to take totalClean and calculate percent change from our two time points.
# But first we need to separate them out into frags and wilds

# Separating into Frags because we need to add up the trials
totalFrag <- totalClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")

# Separating into wilds 
totalWild <- totalClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)

# Combining the two back together
totalCombo <- bind_rows(totalFrag,totalWild)

# Now we are separating out each bracket of time points so we can calculate Percent change from T0 to T3 and T0 to T6.

totalT3 <- totalCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3")) %>%
  select(c('timepoint','species','id','treatment', "SA"))

totalT6 <- totalCombo %>%
  dplyr::filter(timepoint %in% c("T0","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))

# Going from long to wide so we can do the calculation
totalChangeT3 <- dcast(totalT3,id+species+treatment~timepoint) 

# Calculating Percent Change
totalChangeT3 <- dplyr::rename(totalChangeT3, "initialSurfaceArea" = "T0", "T3SurfaceArea" = "T3")
totalChangeT3
totalChangeT3$difference <- (totalChangeT3$T3SurfaceArea-totalChangeT3$initialSurfaceArea)
totalChangeT3$T3percentChange <- ((totalChangeT3$T3SurfaceArea-totalChangeT3$initialSurfaceArea)/totalChangeT3$initialSurfaceArea)*100

# Going from long to wide so we can do the calculation
totalChangeT6 <- dcast(totalT6, id+species+treatment~timepoint) 

# Calculating Percent Change
totalChangeT6 <- dplyr::rename(totalChangeT6, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
totalChangeT6
totalChangeT6$difference <- (totalChangeT6$finalSurfaceArea-totalChangeT6$initialSurfaceArea)
totalChangeT6$T6percentChange <- ((totalChangeT6$finalSurfaceArea-totalChangeT6$initialSurfaceArea)/totalChangeT6$initialSurfaceArea)*100

# combining the two datasets again 
totalChange <- bind_rows(totalChangeT3, totalChangeT6)

# Selecting so we can melt the data back from wide to long again
totalChange <- totalChange %>% 
  dplyr::select(id,species, treatment,T3percentChange,T6percentChange)
totalChangeMelt1 <- melt(totalChange, id = c(1,2,3))

# Cleaning up the data after melting it
totalChangeMelt <- totalChangeMelt1 %>% 
  na.omit() %>% 
  tidyr::separate(variable, into = c("timepoint", "remove" ), sep = 2) %>% 
  dplyr::rename("percentChange" = "value") %>% 
  dplyr::select(id,species, treatment,timepoint,percentChange)

# Renaming the dataframe 
totalChange <- totalChangeMelt


```


# Normality testing and Data Transformation
***
## Here we are testing for normality of the raw SA data, and if not normally distributed, we are transforming the data to see if that will fit a normal distribution.

```{r, normalityTesting, include = TRUE}
# totalPerm <- adonis(formula = SA ~ treatment*timepoint*species, data = totalSA, method = "euclidian", permutations = 9999 )

#Testing normality of totalSA and totalChangeSA

hist(totalSA$SA)

shapiro.test(totalSA$SA)

#Significant so Data does not fit normal distribution
#log transform data

totalSA$logSA <- log(totalSA$SA)

#Re-testing normality
hist(totalSA$logSA)

shapiro.test(totalSA$logSA)



```
# Statistical Analyses
***
## With our newly formed dataset and our transformed data, we are now going to perform some Repeated Measures Analyses of variance using the `lmer()` function from the package `lme4`. We are testing our logtransformed raw SA data as well as our percent change data with **treatment** and **timepoint** as factors and **id** as a random effect.


```{r, repeated measures, include = TRUE}


logTotalLM <- lmer(logSA ~ species*treatment*timepoint + (1 | id), totalSA)
logTotalLManova <- ranova(logTotalLM)
logTotalLManova



eta_sq(logTotalLM, partial = TRUE)
r.squaredGLMM(logTotalLM)
emmeans(logTotalLM, list(pairwise ~ timepoint), adjust = 'bonferroni')



# 
# percentChangeAOV <- aov(percentChange ~ treatment, totalChangeSA )
# summary(percentChangeAOV)



changeLM <- lmer(percentChange ~ species*treatment*timepoint + (1 | id), totalChange)
changeLManova <- anova(changeLM)
changeLManova



```


```{r, visualization, include = TRUE}
se <- function(x) sqrt(var(x) / length(x))

totalSAse <- totalSA %>%
  dplyr::select(timepoint, species, treatment, SA, logSA) %>% 
  dplyr::group_by(timepoint, species, treatment) %>% 
  dplyr::summarise(
    meanSA = mean(SA),
    sd = sd(SA),
    n = n(),
    se = sd/sqrt(n))

totalChangeSe <- totalChange %>% 
dplyr::select(timepoint, species, treatment,percentChange) %>% 
  dplyr::group_by(timepoint, species, treatment) %>% 
  dplyr::summarise(
    meanPC = mean(percentChange),
    sd = sd(percentChange),
    n = n(),
    se = sd/sqrt(n))

# totalSA$se <- se(totalSA$SA)
# totalChange$se <-se(totalChange$percentChange)

SApercentChange.1 <- ggplot(totalChangeSe, aes(x = treatment, y = meanPC, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~species)+
    ggtitle("Surface Area Percent Growth of D. labyrinthiformis, P. strigosa, and P. clivosa Over Six Months") + scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Whole"))+scale_fill_manual(name="Time Point",values = c( "#EAA65D","#0D3B66"))+    
    xlab("Treatment") + ylab("Percent Change (%)")+
  geom_errorbar( aes(x=treatment, ymin=meanPC-se, ymax=meanPC+se), position = position_dodge()) +
    ggtitle("Mean Percent Change in Surface Area of D. labyrinthiformis, P. strigosa, and P. clivosa Over Six Months") + scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Whole"))+scale_fill_manual(name="Time Point",values = c("#0D3B66","#EAA65D"))+    
    xlab("Treatment") + ylab("Percent Change (%)")
SApercentChange.1

SApercentChange<- SApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_blank(), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

SApercentChange

ggsave("../figures/SApercentChange.png", plot = SApercentChange, height = 15, width = 20, units = 'in',dpi = 600)

```

```{r, visualization, include = TRUE}



SAgrowth.1 <- ggplot(data=totalSAse, aes(x = treatment, y = meanSA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  geom_errorbar( aes(x=treatment, ymin=meanSA-se, ymax=meanSA+se), position = position_dodge())+ 
  facet_wrap(~species)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Mean Surface Area Growth of D. labyrinthiformis, P. strigosa, and P. clivosa Over Six Months") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Whole"))+scale_fill_manual(values = c("#82CDED", "#2B4F2E", "#EAA65D"))

SAgrowth.1

SAgrowth <- SAgrowth.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


SAgrowth

ggsave("../figures/SAgrowth.png", plot = SAgrowth, height = 15, width = 20, units = 'in',dpi = 600)

```


```{r, visualization, include = TRUE}


# meanPercentChange.1 <- ggplot(meanDataSet, aes(x = timepoint, y = meanPercentChange, fill = treatment)) +
#   geom_bar(stat = "identity", position = position_dodge())+
#   facet_wrap(~species)+
#     ggtitle("Mean Surface Area Percent Growth of D. labyrinthiformis, P. clivosa, and P. strigosa Over Six Months") +scale_fill_manual(name="Time Point",labels=c("frag" = "Frag", "wild" = "Whole"),values = c( "#EAA65D","#0D3B66"))+    
#     xlab("Time Point") + ylab("Percent Change (%)")
# meanPercentChange.1
# 
# meanPercentChange<- meanPercentChange.1 + theme(
#   panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
#   panel.background = element_rect(fill = '#F5F5F5'),
#   plot.title = element_text(hjust = 0.5), 
#   axis.line = element_line(colour = "black"), 
#   axis.ticks = element_line(color="black"), 
#   text = element_text(size = 35, color="black"), 
#   axis.text.x=element_text(size = 31, color="black"), 
#   axis.text.y=element_text(size = 31, color="black"),
#   legend.position = "right")
# 
# meanPercentChange
# 
# ggsave("../figures/meanPercentChange.png", plot = meanPercentChange, height = 15, width = 30, units = 'in',dpi = 600)

```
