---
title: "Surface Area Analysis"
author: "Ian Combs -- icombs@mote.org"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
knitr::opts_knit$set(root.dir = "../data")
options(width = 88)
library(magrittr)
```





### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

<!-- this is where the DOI would go  [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3675991.svg)](https://doi.org/10.5281/zenodo.3675991)
-->


#### [GitHub repository](https://github.com/icombs2017/fragGrowthExperiment.git){target="_blank"}
###

***
This is the working analysis pipeline to analyze data generated from 3D models of micro-fragmented coral colonies and larger frags to assess differences in growth rates between microfragments and natural growth. 

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 


```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "googlesheets4", "dplyr", "officer","reshape2", "stringr", "flextable", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "vegan", "googledrive", "gdata", "readxl", "DescTools","patchwork", "FSA", "rstatix", "tidyverse")

#IF YOU WANT TO RUN PAIRWISE PERMANOVA REMEMBER TO INSTALL PAIRWISE ADONIS

```


# Importing Data
***
## We are downloading this dataset from our GoogleDrive folder. We will be using the package `googledrive`. Each GoogleDrive file has a unique ID that does not change throughout the lifespan of the document, even if the file name is changed. This ID is housed in the file's URL. 
Example: docs.google.com/spreadsheets/d/FILE_ID_GOES_HERE/other_information/. Below you will copy and paste that into the function `drive_download` within `as_id`. This will save the file locally in the specified path (in this case our data folder) and you will import the folder as you normally would. Downloading it this way decreases the potential human error when downloading, moving folders, renaming, saving etc and ensures that the most up to date file is being utilized. 
```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

dlsa <- read_excel("../data/saDataset.xlsx", sheet = 3)

dlsa


```
## Now I am just tidying up my dataset, I am combining two columns, **species** and **genotype** into a column **geno** using the `paste0` function, cleaning up the raw dataset to only include my species of concern using multiple functions from the package `dplyr`, and changing the class of some variables to factors using the function `as.factor()` for easier manipulation down the road.
```{r, dataManipulation, include = TRUE}


dlsaClean <- dlsa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




dlsaClean$timepoint <- as.factor(dlsaClean$timepoint)

dlsaClean$species <- as.factor(dlsaClean$species)

dlsaClean$id <- as.factor(dlsaClean$id)

dlsaClean$trial <- as.factor(dlsaClean$trial)

dlsaClean$treatment <- as.factor(dlsaClean$treatment)



head(dlsaClean)



```
### Here we are breaking the data apart into frags and wilds, combinging the frags from each trial, putting the data back together, and then calculating percent change of surface area between the first and last timepoint. We are using a number of functions from the package `dplyr` as well as the function `dcast()` to recombine data from the package `reshape2` 
```{r, calculatingPercentChange, include = TRUE }


dlsaFrag <- dlsaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


dlsaWild <- dlsaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


dlsaCombo <- bind_rows(dlsaFrag,dlsaWild)

dlsaFinal <- dlsaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



dlsaChange <- dcast(dlsaFinal, id+treatment~timepoint) 


dlsaChange <- dplyr::rename(dlsaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
dlsaChange
dlsaChange$difference <- (dlsaChange$finalSurfaceArea-dlsaChange$initialSurfaceArea)
dlsaChange$percentChange <- ((dlsaChange$finalSurfaceArea-dlsaChange$initialSurfaceArea)/dlsaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



dlSApercentChange.1 <- ggplot(data=dlsaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) +  scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E")) +     
    xlab("ID") + ylab("Percent Change (%)") + # Set axis labels
    ggtitle("D. lab") +     # Set title
    theme_bw()

dlSApercentChange <- dlSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

dlSApercentChange1

ggsave("../figures/dlPercentSAgrowth.png", plot = dlSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



dlSAgrowthGenotype.1 <- ggplot(data=dlsaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of DL01, DL02, and DL03") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#3BB0E3", "#eba862", "#12A528"))

dlSAgrowthGenotype.1

dlSAgrowthGenotype <- dlSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


dlSAgrowthGenotype

ggsave("../figures/dlSAgrowth.png", plot = dlSAgrowthGenotype, height = 5, width = 9, units = 'in',dpi = 600)



```

```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

pcsa <- read_excel("../data/saDataset.xlsx", sheet = 4)

pcsa


```

```{r, dataManipulation, include = TRUE}


pcsaClean <- pcsa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




pcsaClean$timepoint <- as.factor(pcsaClean$timepoint)

pcsaClean$species <- as.factor(pcsaClean$species)

pcsaClean$id <- as.factor(pcsaClean$id)

pcsaClean$trial <- as.factor(pcsaClean$trial)

pcsaClean$treatment <- as.factor(pcsaClean$treatment)



head(pcsaClean)



```

```{r, calculatingPercentChange, include = TRUE }


pcsaFrag <- pcsaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


pcsaWild <- pcsaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


pcsaCombo <- bind_rows(pcsaFrag,pcsaWild)

pcsaFinal <- pcsaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



pcsaChange <- dcast(pcsaFinal, id+treatment~timepoint) 


pcsaChange <- dplyr::rename(pcsaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
pcsaChange
pcsaChange$difference <- (pcsaChange$finalSurfaceArea-pcsaChange$initialSurfaceArea)
pcsaChange$percentChange <- ((pcsaChange$finalSurfaceArea-pcsaChange$initialSurfaceArea)/pcsaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



pcSApercentChange.1 <- ggplot(data=pcsaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) + scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E")) +   
    xlab("ID") + ylab("Percent Change (%)") + # Set axis labels
    ggtitle("P. cli") +     # Set title
    theme_bw()

pcSApercentChange <- pcSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

pcSApercentChange

ggsave("../figures/pcPercentSAgrowth.png", plot = pcSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



pcSAgrowthGenotype.1 <- ggplot(data=pcsaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of PC13, PC15, and PC16") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#3BB0E3", "#eba862", "#80f191"))

pcSAgrowthGenotype.1

pcSAgrowthGenotype <- pcSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


pcSAgrowthGenotype

ggsave("../figures/pcSAgrowth.png", plot = pcSAgrowthGenotype, height = 5, width = 9, units = 'in',dpi = 600)



```

```{r, loadingData, include = TRUE}


drive_download(
  as_id("1C1deEPnb07YcvbEngx_OV3Xl-C740sBefoaum154OyY"),
  path = "../data/saDataset.xlsx", 
  overwrite = TRUE)

pssa <- read_excel("../data/saDataset.xlsx", sheet = 5)

pssa


```

```{r, dataManipulation, include = TRUE}


pssaClean <- pssa %>% 
  select("timepoint", "species", "genotype","id","trial","treatment","SA")




pssaClean$timepoint <- as.factor(pssaClean$timepoint)

pssaClean$species <- as.factor(pssaClean$species)

pssaClean$id <- as.factor(pssaClean$id)

pssaClean$trial <- as.factor(pssaClean$trial)

pssaClean$treatment <- as.factor(pssaClean$treatment)



head(pssaClean)



```

```{r, calculatingPercentChange, include = TRUE }


pssaFrag <- pssaClean %>%
  select(timepoint, species, id, treatment, SA) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(timepoint, species, id, treatment) %>%
  dplyr::summarise(sum(SA)) %>% 
  dplyr::rename("SA" = "sum(SA)")


pssaWild <- pssaClean %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(timepoint, species, id, treatment, SA)


pssaCombo <- bind_rows(pssaFrag,pssaWild)

pssaFinal <- pssaCombo %>%
  dplyr::filter(timepoint %in% c("T0","T3","T6")) %>%
  select(c('timepoint','species','id','treatment', "SA"))



pssaChange <- dcast(pssaFinal, id+treatment~timepoint) 


pssaChange <- dplyr::rename(pssaChange, "initialSurfaceArea" = "T0", "finalSurfaceArea" = "T6")
pssaChange
pssaChange$difference <- (pssaChange$finalSurfaceArea-pssaChange$initialSurfaceArea)
pssaChange$percentChange <- ((pssaChange$finalSurfaceArea-pssaChange$initialSurfaceArea)/pssaChange$initialSurfaceArea)*100

 

```

```{r, visualization, include = TRUE}



psSApercentChange.1 <- ggplot(data=pssaChange, aes(x=id, y=percentChange, fill=treatment)) + 
    geom_bar(colour="black", stat="identity",
             position=position_dodge(),
             size=.3) +scale_fill_manual(name="Treatment",labels=c('Frag', 'Wild'),values = c("#82CDED", "#2B4F2E"))+      
    xlab("ID") + ylab("Percent Change (%)") + 
    ggtitle("P. str") +     
    theme_bw()

psSApercentChange <- psSApercentChange.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")

psSApercentChange

ggsave("../figures/psPercentSAgrowth.png", plot = psSApercentChange, height = 5, width = 9, units = 'in',dpi = 600)


```

```{r, visualization, include = TRUE}



psSAgrowthGenotype.1 <- ggplot(data=pssaFinal, aes(x = treatment, y = SA, fill = timepoint)) +
  geom_bar(stat = "identity", position = position_dodge())+
  facet_wrap(~id)+
  scale_fill_hue(name="Time Point") +  
  labs(x = "Treatment", y = bquote("Surface Area" ~ (cm^2)),fill='Time Point') + # Set axis labels
    ggtitle("Surface Area Growth of PS07, PS11, PS12, PS13, and PS14") + scale_fill_discrete(labels=c('Initial','T3','Final'))+scale_x_discrete(labels=c("frag" = "Frag", "wild" = "Wild"))+scale_fill_manual(values = c("#82CDED", "#2B4F2E", "#EAA65D"))

psSAgrowthGenotype.1

psSAgrowthGenotype <- psSAgrowthGenotype.1 + theme(
  panel.grid.major = element_line(size = 0.5, linetype = 'solid', colour = "white"),
  panel.background = element_rect(fill = '#F5F5F5'),
  plot.title = element_text(hjust = 0.5), 
  axis.line = element_line(colour = "black"), 
  axis.ticks = element_line(color="black"), 
  text = element_text(size = 22, color="black"), 
  axis.text.x=element_text(size = 18, color="black"), 
  axis.text.y=element_text(size = 18, color="black"),
  legend.position = "right")


psSAgrowthGenotype

ggsave("../figures/psSAgrowth.png", plot = psSAgrowthGenotype, height = 10, width = 15, units = 'in',dpi = 600)



```

```{r}
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

percentChange <- ggarrange(dlSApercentChange, pcSApercentChange, psSApercentChange,
                    labels = c("A", "B", "C"),
                    ncol = 2, nrow = 2, common.legend = TRUE, legend = "right")
percentChange <- annotate_figure(
  percentChange,
  top = text_grob("Surface Area Percent Change",
                  color = "black", face = "bold", size = 24))


# ggarrange(
#   dlSApercentChange,                
#   ggarrange(pcSApercentChange, psSApercentChange, ncol = 2, labels = c("B", "C")), 
#   nrow = 2, 
#   labels = "A"      
#   ) 
percentChange

ggsave("../figures/percentChange.png", plot = percentChange, height = 15, width = 20, units = 'in',dpi = 600)

```
```{r}

percentChange.1<-dlSApercentChange+ pcSApercentChange+psSApercentChange+plot_layout(nrow = 2, byrow = FALSE)+plot_annotation(title = 'Surface Area Percent Change',
                  theme = theme(plot.title = element_text(size = 24)))

percentChange.1

ggsave("../figures/percentChange.1.png", plot = percentChange.1, height = 15, width = 20, units = 'in',dpi = 600)
```

Now we will run a paired t-test to determine if the percent change in surface area of the frags varies from the wilds. First, we need to assess normality via the shapiro-wilk test for normal distribution and the Levene's Test to test for homogeneity of variance. If neither is significant, we will conduct a t-test.
```{r}
dlsaFragTest <- dlsaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

dlsaWildTest <- dlsaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

dlsaTest <- left_join(dlsaFragTest, dlsaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(fragpercentChange = percentChange.x) %>%
  dplyr::rename(wildpercentChange = percentChange.y)
```


```{r}
shapiro.test(dlsaChange$percentChange) 
  
x=dlsaTest$fragpercentChange
y=dlsaTest$wildpercentChange

# LeveneTest(dlsaCombo$surfaceArea, group = dlsaCombo$treatment)

t.test(x, y, paired = TRUE, alternative = "two.sided")



```
```{r}
pcsaFragTest <- pcsaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

pcsaWildTest <- pcsaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

pcsaTest <- left_join(pcsaFragTest, pcsaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(pcfragpercentChange = percentChange.x) %>%
  dplyr::rename(pcwildpercentChange = percentChange.y)
```


```{r}
shapiro.test(pcsaChange$percentChange) 
  
x=pcsaTest$pcfragpercentChange
y=pcsaTest$pcwildpercentChange

# LeveneTest(pcsaCombo$surfaceArea, group = pcsaCombo$treatment)

t.test(x, y, paired = TRUE, alternative = "two.sided")

```
```{r}
pssaFragTest <- pssaChange %>%
  select(id, treatment, percentChange) %>%
  dplyr::filter(treatment %in% 'frag') %>%
  dplyr::group_by(id,treatment)

pssaWildTest <- pssaChange %>%
  dplyr::filter(treatment %in% 'wild') %>%
  select(id, treatment, percentChange)

pssaTest <- left_join(pssaFragTest, pssaWildTest, by = c("id"), copy = FALSE) %>%
  dplyr::rename(psfragpercentChange = percentChange.x) %>%
  dplyr::rename(pswildpercentChange = percentChange.y)
```


```{r}
shapiro.test(pssaChange$percentChange) 
  
x=pssaTest$psfragpercentChange
y=pssaTest$pswildpercentChange

# LeveneTest(pcsaCombo$surfaceArea, group = pcsaCombo$treatment)

t.test(x, y, paired = TRUE, alternative = "two.sided")

```
# Here I am finding the mean percent change values
```{r}
dlWildMeanPercentChange <- mean(dlsaWildTest$percentChange)

dlWildMeanPercentChange

dlFragMeanPercentChange <- mean(dlsaFragTest$percentChange)

dlFragMeanPercentChange

pcWildMeanPercentChange <- mean(pcsaWildTest$percentChange)

pcWildMeanPercentChange

pcFragMeanPercentChange <- mean(pcsaFragTest$percentChange)

pcFragMeanPercentChange

psWildMeanPercentChange <- mean(pssaWildTest$percentChange)

psWildMeanPercentChange

psFragMeanPercentChange <- mean(pssaFragTest$percentChange)

psFragMeanPercentChange
```


we are going to combine all of the data we've parsed out above

```{r, stats, include = TRUE}

totalChangeSA <-  bind_rows(dlsaChange, pcsaChange, pssaChange)
totalSA <- bind_rows(dlsaFinal, pcsaFinal, pssaFinal)

totalPerm <- adonis(formula = SA ~ treatment*timepoint*species, data = totalSA, method = "euclidian", permutations = 9999 )


#log transform data

totalSA$logSA <- log(totalSA$SA)

SAtTest <- t.test(logSA  ~ treatment, data = totalSA)

SAaov <- aov(SA ~ )

```

```{r, repeated measures, include = TRUE}


totalSA %>%
  group_by(timepoint) %>%
  identify_outliers(SA)


totalSA %>%
  group_by(treatment, timepoint) %>%
  shapiro_test(logSA)

ggqqplot(totalSA, "logSA", ggtheme = theme_bw()) +
  facet_grid(timepoint ~ treatment, labeller = "label_both")



one.way <- totalSA %>%
group_by(timepoint) %>%
anova_test(dv = logSA, wid = id, within = treatment) %>%
get_anova_table() %>%
adjust_pvalue(method = "bonferroni")
one.way

pwc2 <- totalSA %>%
group_by(treatment) %>%
pairwise_t_test(
logSA ~ timepoint, paired = TRUE,
p.adjust.method = "bonferroni"
)
data.frame(pwc2)





```









