---
title: "Surface Growth Analysis"
author: "Kamryn Herbison -- ENTER EMAIL HERE"
output:
  html_document:
    theme: flatly
    code_folding: show
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_doctument:
      toc: yes
      toc_depth: 3
---
```{r, setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, fig.align = 'left')
library(magrittr)
```

### version: `r Sys.Date() %>% format(format="%B %d, %Y")`

[![DOI](https://zenodo.org/badge/255665897.svg)](https://zenodo.org/badge/latestdoi/255665897)



#### [GitHub repository](https://github.com/icombs2017/analysisOf3dModels.git){target="_blank"}


###
***
This is the analysis pipeline to analyze data generated from 3D models of fate tracked *Montastraea cavernosa* infected with stony coral tissue loss disease. For the manuscript **Combs IR, Studivan MS, Voss JD. (2020) 

***

### All analyses performed with R version `r getRversion()`


# Basic setup of R environment
***
## Loading required packages
For the following analyses we will require the use of a number of different R packages. Most of which can be sourced from CRAN, but some must be downloaded from GitHub. We can use the following code to load in the packages and install any packages not previously installed in the R console. 

```{r,packages, include = TRUE, message = FALSE, warning = FALSE, results = 'hide'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load("ggplot2", "FSA", "officer", "gridExtra", "ggpubr", "Rmisc", "rcompanion", "RColorBrewer", "vegan", "nparcomp", "RVAideMemoire", "pairwiseAdonis", "PMCMR", "PMCMRplus", "patchwork", "magrittr","reshape2", "stringr", "dplyr", "flextable", "permuco")
pacman::p_load_gh("pmartinezarbizu/pairwiseAdonis/pairwiseAdonis")
```
<br><br>


```{r}




```










